\documentclass[FrontPage]{grattan}
% Comments are deployed by the % sign; everything after % is ignored by the compiler.
% Please do not put comments before \documentclass as these are reserved for TeX directives.
% add_to_dictionary: UAC ATAR ATARs UTAS disenrol disenrolled disenrolling disenrolment EFTSL misclassify timeframe timeframes CHESSN CHESSNs TEQSA QILT ANOP DEET somethings RMIT SurveyMonkey IEO GPA LSAY TAFE UES

\usepackage{longtable}
\addbibresource{bib/HigherEd-Grattan.bib}

% release: true

\author{Andrew Norton \& Ittima Cherastidtham}
\title{Dropping out: the benefits and costs of trying university}

\GrattanReportNumber{2018-07}

% editorial_author_only: Paul Austin
\acknowledgements{%
This report was written by Andrew Norton, Grattan Institute Higher Education Program Director and Ittima Cherastidtham, Higher Education Program Fellow. Will Mackey provided extensive research assistance and made a substantial contribution to the report. James Ha also provided research assistance. Paul Austin provided substantial editing assistance.

We would like to thank members of Grattan Institute's Higher Education Reference Group, Buly Cardak, Marcia Devlin, Toni Downes, Richard James, Kevin Jameson, Ian Li, Mike Evans and several others for their advice on this report and its background paper. 

The opinions in this report are those of the authors and do not necessarily represent the views of Grattan Institute's founding members, affiliates, individual board members, reference group members or reviewers.
Any remaining errors or omissions are the responsibility of the authors.

Grattan Institute is an independent think-tank focused on Australian public policy.
Our work is independent, practical and rigorous.
We aim to improve policy outcomes by engaging with both decision-makers and the community.

For further information on the Institute's programs, or to join our mailing list, please go to: \textcolor{blue}{\url{http://www.grattan.edu.au/}}.


{\footnotesize
This report may be cited as:
Norton, A., Cherastidtham, I., and Mackey, W\@. (2018). \emph{\mytitle}. Grattan Institute.

ISBN: 978-0-6482307-7-9             %%updated and correct at 24-04-18

All material published or otherwise created by Grattan Institute is licensed under a Creative Commons Attribution-NonCommercial-ShareAlike 3.0 Unported License\par
}
}


\begin{document}
    \input{report/overview}
    \input{report/recommendations}
    
    \contentspage
    
    \listoffigures
    
    \input{report/chap1}
    \input{report/chap2}
    \input{report/chap3}
    \input{report/chap4}
    \input{report/chap5}
    \input{report/chap6}
    \input{report/chap7}
    \input{report/chap8}

\appendix
    \input{report/appendix_a}
    \input{report/appendix_b}
    \input{report/glossary}     % glossary is current as of 26/04/18


\printbibliography


\end{document}


